Using Anchor Points to Define and Transfer Spatial Regions Based on Context

### 2012-05-15 code/scenarios/cdsr/trunk/papers/QR2012/main-QR12-cdsr.tex

\documentclass[letterpaper]{article}
\usepackage{aaai}

@@

\maketitle
\begin{abstract}
In order to collaborate with people in the real world, AI systems must be able to represent and reason about spatial regions in human environments. Consider the command \emph{``go to the front of the classroom''}. The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone. Instead it is defined by its functional use, implied by nearby objects and their configuration. In this paper, we define such areas as \textit{context-dependent spatial regions} and present a system able to learn them by combining qualitative spatial representations, semantic labels, and analogy. The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world. It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.  From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms. To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
\end{abstract}

@@

This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs. Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) corresponding to sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically and geometrically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs. Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

@@

In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, the QSRs between them and any groups that have been identified. The base also contains a labeled CDSR of the type sought in target. The result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target. We use these to transfer the CDSR from base to target (i.e. recognizing the CDSR in the target) as follows. First, we identify the CDSR of the sought type in the base and use SME's pragmatic constraints to ensure that the entities referred to its anchor points participate in the mapping. To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{boundarySegment} statements mentioning the base CDSR. The second and third arguments of these candidate inferences are anchor points in the target environment. We use these to define the boundary of the CDSR in the target.

>>>>

\documentclass[letterpaper]{article}
\usepackage{aaai}
\nocopyright

@@

\maketitle
\begin{abstract}
In order to collaborate with people in the real world, AI systems must be able to represent and reason about spatial regions in human environments. Consider the command \emph{``go to the front of the classroom''}. The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone. Instead it is defined by its functional use, implied by nearby objects and their configuration. In this paper, we define such areas as \textit{context-dependent spatial regions} and present a system able to learn them by combining qualitative spatial representations, semantic labels, and analogy. The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world. It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.  From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms by transferring the necessary anchor points. To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
\end{abstract}

@@

This paper presents an artificial cognitive system (specifically a mobile robot) able to represent and reason about CDSRs. Central to our approach is the use of \emph{anchor points}, symbolic expressions which link conceptual entities (e.g., CDSRs) to perceived entities (e.g., objects in the environment). Our approach is founded on two assumptions. The first assumption is that CDSRs can be defined using \emph{qualitative spatial representations} (QSRs) corresponding to sensor data of the system~\cite{Cohn:2001}. The second assumption is that semantically and geometrically similar areas (e.g. two different classrooms) will feature similar CDSRs, and that these similarities can be recognised through \emph{analogy}. The rest of the paper is structured following these assumptions. Section~\ref{sec:qsr-gen} describes how we generate QSRs from sensor data taken from an existing, state-of-the-art, cognitive system and use these to define CDSRs with anchor points. Section~\ref{sec:analogy} then describes how we use the structure-mapping model of analogy \cite{Gentner1983a} to transfer a CDSR from a labelled example to a new situation. Section~\ref{sec:example} presents a worked example of the entire process, and Section~\ref{sec:evaluation} evaluates our system in comparison to data from human subjects performing the same task.

@@

In our system, the base and target representations consist of the entities Dora has perceived in two different rooms, the QSRs between them and any groups that have been identified. The base also contains a labeled CDSR of the type sought in target. The result of running SME on these representations is a set of correspondences between the base and target, and a set of candidate inferences about the target. We use these to transfer the CDSR from base to target (i.e. recognizing the CDSR in the target) as follows. First, we identify the CDSR of the sought type in the base and use SME's pragmatic constraints to ensure that the entities referred to by its anchor points participate in the mapping. For example, if the CDSR in the base included the anchor point \fw{(YMaxXMostFn Desk1)}, then we would include a constraint requiring \fw{Desk1} participate in the mapping. To transfer the CDSR to the target, we collect the candidate inferences that result from \fw{boundarySegment} statements mentioning the base CDSR. The second and third arguments of these candidate inferences are anchor points in the target environment. We use these anchor points to define the boundary of the CDSR in the target.

### 2012-05-15 code/scenarios/cdsr/trunk/papers/QR2012/qsr-extraction.tex

We generate the strengths of these spatial relations as follows. First we compute the maximum distance $d_{max}$ between any two points in the room, this value is used to normalize the distances between objects. Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid. This results in the coordinates of the all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.  We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance $d$ between the landmark's centroid and the adjacent object's location, and (b) the inner angle $\theta$ between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location. These two spatial components are integrated to compute the strength $s$ of a given relationship using Equation~\ref{eq:spatial}. Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region.

>>>>

We generate the strengths of these spatial relations as follows. First we compute the maximum distance $d_{max}$ between any two points in the room, this value is used to normalize the distances between objects. Next, taking each object in turn to be the landmark, we translate the origin of the room to the landmark's centroid. This results in the coordinates of all the other objects in the room being translated into a frame of reference whose origin is the centroid of the landmark.  We then compute the strength of each of the 8 spatial relations between the landmark and each of the objects adjacent to it by calculating: (a) the distance $d$ between the landmark's centroid and the adjacent object's location, and (b) the inner angle $\theta$ between the direction vector of the relation and the vector from the origin (the landmark's centroid) to the neighbour's location. These two spatial components are integrated to compute the strength $s$ of a given relationship using Equation~\ref{eq:spatial}. Figure \ref{fig:spr} provides a visualization of a spatial relationship across a region.
